\MakeUppercase{Linear versus centred Colouring Numbers}\thanks{This research was partly funded by NSERC.}

### 2021-10-05 lin-vs-cen.tex
Pfffffft

\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird conditions.}

@@

  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=O(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.  To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.  (The remainder of the proof does not depend on the values of $k$ or $C$.)

@@

>>>>

\todo[inline]{I can't yet prove the full result, which is why the next lemma has some weird condition involving a newly introduced parameter $r$.}

@@

  Returning to the proof of \cref{pseudogrid_lower_bound}, we set $\epsilon=1/2\alpha=\Theta(1/r\beta^2)$ so that \cref{all_frequent} guarantees that $H'$ contains a $k/2\times k/2$ pseudogrid in which each colour $c\in\{1,\ldots,C\}$ either does not appear in $H'$ or is frequent in $H'$.  To avoid overloading notation, we now assume that there is no infrequent colour, i.e., each colour $c\in\{1,\ldots,C\}$ has multiplicity at least $\alpha$ in $H$.  (The remainder of the proof does not depend on the values of $k$ or $C$.)

@@

To finish the proof, we need some lemma which states that highly subdivided graphs don't have linear colouring numbers that are much smaller than mildly subdivided graphs.  It would be enough to prove the following.

A \emph{refinement} $G'$ of a graph $G$ is a subdivision of $G$ in which each subdivision vertex is adjacent only to degree-$2$ vertices, i.e., $\deg_{G'}(v)=2$ for each $v\in N_{G'}(s)$ and each $s\in V(G')\setminus V(G)$.

\begin{lem}\label{refinement_bound}
  For any refinement $G'$ of any graph $G$, $\chilin(G') = \Omega(\chilin(G))$.
\end{lem}

In general, we know that if $G'$ is an $\ell$-subdivision of $G$ then
\[  
    \log\ell \le \chilin(G') \le \chilin(G) + \log\ell \enspace .
\] 
The lower bound comes from the fact that a path of length $\ell$ has a linear chromatic number at least $\log\ell$. The upper bound comes by introducing $\log\ell$ additional colours that are used to colour the paths in $G'-G$.  Of course, the upper bound holds even when $G'$ is a $\le\!\!\ell$-subdivision of $G$ and the lower bound holds as long as at least one edge of $G$ is subdivided $\ell$ or more times.

Perhaps \cref{refinement_bound} is true even if we replace ``refinement'' with ``subdivision''.  As a test case, one could start with the $1$-subdivision $K^{\onesub}_k$ of the $k$-clique $K_k$.  The following result is certainly true, as can be seen from a treedepth decomposition:

\begin{lem}
  For any subdivision $G'$ of any graph $G$, $\chicen(G') \ge \chicen(G)$.
\end{lem}
